\section{One-factor Hull-White Model}
The one-factor Hull-White model \cite{Hull1990} is an extension of the Vasicek model such that the parameters are allowed to be time-dependent, i.e.
\begin{equation}
dr_t = \kappa_t \left( \theta_t - r_t \right) dt + \sigma_t dW_t. \label{eq:HullWhiteSDE}
\end{equation}
where the SDE parameters have the same interpretation as in the Vasicek model, and, again, $W_t$ denotes a one-dimensional Wiener process under the risk-neutral $Q$-measure.

For the Hull-White model to be consistent with the initial yield curve, $\theta_t$ is given by \cite{Andersen2010}
\begin{equation}
\theta_t = \frac{1}{\kappa_t} \frac{\partial f(0,t)}{\partial t} + f(0,t) + \frac{1}{\kappa_t} \int_0^t \e^{-2\int_u^t \kappa_v dv} \sigma_u^2 du.
\end{equation}

\subsection{Andersen-Piterbarg transformation}
Now we define a pseudo short rate as $x_t = r_t - f(0,t)$ such that
\begin{equation}
dx_t = \left( y_t - \kappa_t x_t \right) dt + \sigma_t dW_t, \;\;\; x_0 = 0,
\label{eq:HullWhiteSDEx}
\end{equation}
where
\begin{equation}
y_t = \int_0^t \e^{-2\int_u^t \kappa_v dv} \sigma_u^2 du,
\end{equation}
thus, the initial yield curve is not used explicitly in the simulation of the pseudo 
short rate, and any numerical issues associated with differentiation of the yield curve are avoided.

The SDE in Eq.~(\ref{eq:HullWhiteSDEx}) can be solved analytically by introducing the function
\begin{equation}
h(t, x_t) = x_t \e^{\int_0^t \kappa_u du},
\end{equation}
using It\^{o}'s lemma
\begin{eqnarray}
dh(t, x_t) &=& \frac{\partial h}{\partial t} dt + \frac{\partial h}{\partial x_t} dx_t + \frac{1}{2}\frac{\partial^2 h}{\partial x_t^2} \left(dx_t\right)^2 \\
&=& \kappa_t h(t, x_t) dt + \e^{\int_0^t \kappa_u du} dx_t \\
&=& y_t \e^{\int_0^t \kappa_u du} dt + \sigma_t \e^{\int_0^t \kappa_u du} dW_t,
\end{eqnarray}
and integrating from $t_1$ to $t_2$
\begin{equation}
h(t_2, r_{t_2}) = h(t_1, r_{t_1}) + \int_{t_1}^{t_2} y_t \e^{\int_0^t \kappa_u du} dt + \int_{t_1}^{t_2} \sigma_t \e^{\int_0^t \kappa_u du} dW_t,
\end{equation}
such that
\begin{equation}
x_{t_2} = x_{t_1} \e^{-\int_{t_1}^{t_2} \kappa_u du}  + \int_{t_1}^{t_2} y_t \e^{-\int_t^{t_2} \kappa_u du} dt + \int_{t_1}^{t_2} \sigma_t \e^{-\int_t^{t_2} \kappa_u du} dW_t.
\end{equation}

As for the Vasicek model, the pseudo short rate process is Gaussian with conditional mean
\begin{equation}
\E_{t_1}\! \left[ x_{t_2} \right] = x_{t_1} \e^{-\int_{t_1}^{t_2} \kappa_u du}  + \int_{t_1}^{t_2} y_t \e^{-\int_t^{t_2} \kappa_u du} dt,
\end{equation}
and, using the It\^{o} isometry, the conditional variance of the process becomes
\begin{eqnarray}
\Var_{t_1}\! \left[ x_{t_2} \right] &=& \E_{t_1}\! \left[ \left( \int_{t_1}^{t_2} \sigma_t \e^{-\int_t^{t_2} \kappa_u du} dW_t \right)^2 \right] \\
&=& \int_{t_1}^{t_2} \sigma_t^2 \e^{-2\int_t^{t_2} \kappa_u du} dt.
\end{eqnarray}

The corresponding pseudo discount factor is written as
\begin{equation}
D_{t} = \e^{I_t},
\end{equation}
where
\begin{eqnarray}
I_t &=& -\int_0^t x_u du \\
&=& - x_{0} \int_0^t \e^{-\int_{0}^{u} \kappa_v dv} du - \int_0^t\int_{0}^{u} y_v \e^{-\int_v^{u} \kappa_s ds} dv du \\
&& - \int_0^t \int_{v}^{t} \sigma_v \e^{-\int_v^{u} \kappa_s ds} du dW_v,
\end{eqnarray}
where the order of integration has been changed in the last term.

The ``integral'' process is likewise Gaussian with conditional mean
\begin{equation}
\E_{t_1}\! \left[ I_{t_{2}} \right] = I_{t_1} - x_{t_1} \int_{t_1}^{t_2} \e^{-\int_{t_1}^{u} \kappa_v dv} du - \int_{t_1}^{t_2} \int_{t_1}^{u} y_v \e^{-\int_v^{u} \kappa_s ds} dv du,
\end{equation}
and, using the It\^{o} isometry, the conditional variance becomes
\begin{eqnarray}
\Var_{t_1}\! \left[ I_{t_{2}} \right] &=& \E_{t_1}\! \left[ \left( \int_{t_1}^{t_2} \int_{v}^{t_2} \sigma_v \e^{-\int_v^{u} \kappa_s ds} du dW_v \right)^2 \right] \\
&=& \int_{t_1}^{t_2} \sigma_v^2 \left( \int_{v}^{t_2} \e^{-\int_v^{u} \kappa_s ds} du \right)^2 dv.
\end{eqnarray}

\textcolor{red}{Compare (A.50) with Andersen \& Piterbarg!}

Again, we can write both $x_t$ and $I_t$ as a sum of a deterministic term ($D$) and a stochastic term ($S$)
\begin{eqnarray}
x_t &=& D_{x_t} + S_{x_t} \\
I_t &=& D_{I_t} + S_{I_t},
\end{eqnarray}
such that the covariance of $x_t$ and $I_t$ becomes
\begin{eqnarray}
\Cov_{t_1}\! \left[ x_{t_2}, I_{t_2} \right] &=& \E_{t_1}\! \left[ S_{x_{t_2}} S_{I_{t_2}} \right] \\
&=& - \int_{t_1}^{t_2} \int_{t_1}^u \sigma_s^2 \e^{-\int_s^{u} \kappa_v dv} \e^{-\int_s^{t_2} \kappa_v dv} ds du.
\end{eqnarray}

\textcolor{red}{Derive (A.54), and compare with Andersen \& Piterbarg!}

\subsubsection{Transformation from pseudo representation}

The initial discount curve and instantaneous forward rate curve are related by
\begin{equation}
P(0,t) = \e^{- \int_0^t f(0,u) du}.
\end{equation}

Using the transformation of the short rate $r_t = x_t + f(0,t)$, we get that
\begin{eqnarray}
P(0,t) &=& \E \left[ \e^{- \int_0^t r_u du} \right] \\
&=& \e^{- \int_0^t f(0,u) du} \E \left[ \e^{- \int_0^t x_u du} \right],
\end{eqnarray}
which implies that
\begin{equation}
\E \left[ \e^{- \int_0^t x_u du} \right] = 1.
\end{equation}

For a given scenario, i.e., $r_t(\omega) = x_t (\omega) + f(0,t)$, we get the 
time $t$ stochastic discount factor as
\begin{equation}
\e^{- \int_0^t r_u (\omega) du} = P(0,t) \e^{- \int_0^t x_u (\omega) du}.
\end{equation}


\subsection{Pelsser transformation}
First we solve Eq.~\ref{eq:HullWhiteSDE} directly
\begin{equation}
r_{t} = r_{0} \e^{-\int_{0}^{t} \kappa_u du}  + \int_{0}^{t} \kappa_u \theta_u \e^{-\int_u^{t} \kappa_v dv} du + \int_{0}^{t} \sigma_u \e^{-\int_u^{t} \kappa_v dv} dW_u.
\end{equation}

Then we introduce the pseudo short rate $z_t = r_t - \alpha_t$ \cite{Pelsser2000}, where 
\begin{equation}
\alpha_t = r_{0} \e^{-\int_{0}^{t} \kappa_u du}  + \int_{0}^{t} \kappa_u \theta_u \e^{-\int_u^{t} \kappa_v dv} du,
\end{equation}
such that
\begin{equation}
dz_t = - \kappa_t z_t dt + \sigma_t dW_t \;\;\; z_0 = 0,
\end{equation}
hence, again the initial yield curve is not used explicitly in the simulation of pseudo short rate.

The SDE can be solved in a similar fashion to the Andersen-Piterbarg transformation such that
\begin{equation}
z_{t_2} = z_{t_1} \e^{-\int_{t_1}^{t_2} \kappa_t dt} + \int_{t_1}^{t_2} \sigma_t \e^{-\int_t^{t_2} \kappa_u du} dW_t.
\end{equation}

The pseudo short rate process is Gaussian with conditional mean
\begin{equation}
\E_{t_1}\! \left[ z_{t_2} \right] = z_{t_1} \e^{-\int_{t_1}^{t_2} \kappa_u du},
\end{equation}
and, using the It\^{o} isometry, the conditional variance of the process becomes
\begin{eqnarray}
\Var_{t_1}\! \left[ z_{t_2} \right] &=& \int_{t_1}^{t_2} \sigma_t^2 \e^{-2\int_t^{t_2} \kappa_u du} dt.
\end{eqnarray}

The corresponding pseudo discount factor is written as
\begin{equation}
D_{t} = \e^{I_t},
\end{equation}
where
\begin{eqnarray}
I_t &=& -\int_0^t z_u du \\
&=& - z_{0} \int_0^t \e^{-\int_{0}^{u} \kappa_v dv} du - \int_0^t \int_{v}^{t} \sigma_v \e^{-\int_v^{u} \kappa_s ds} du dW_v,
\end{eqnarray}
where the order of integration has been changed in the last term.

The discount factor process is likewise Gaussian with conditional mean
\begin{equation}
\E_{t_1}\! \left[ I_{t_{2}} \right] = I_{t_1} - z_{t_1} \int_{t_1}^{t_2} \e^{-\int_{t_1}^{u} \kappa_v dv} du,
\end{equation}
and, using the It\^{o} isometry, the conditional variance becomes
\begin{eqnarray}
\Var_{t_1}\! \left[ I_{t_{2}} \right] &=& \int_{t_1}^{t_2} \sigma_v^2 \left( \int_{v}^{t_2} \e^{-\int_v^{u} \kappa_s ds} du \right)^2 dv.
\end{eqnarray}

Again, we can write both $z_t$ and $I_t$ as a sum of a deterministic term ($D$) and a stochastic term ($S$)
\begin{eqnarray}
z_t &=& D_{z_t} + S_{z_t} \\
I_t &=& D_{I_t} + S_{I_t},
\end{eqnarray}
such that the covariance of $z_t$ and $I_t$ becomes
\begin{eqnarray}
\Cov_{t_1}\! \left[ z_{t_2}, I_{t_2} \right] &=& \E_{t_1}\! \left[ S_{z_{t_2}} S_{I_{t_2}} \right] \\
&=& - \int_{t_1}^{t_2} \int_{t_1}^u \sigma_s^2 \e^{-\int_s^{u} \kappa_v dv} \e^{-\int_s^{t_2} \kappa_v dv} ds du.
\end{eqnarray}

\subsubsection{Transformation from pseudo representation}

Using the transformation of the short rate $r_t = z_t + \alpha_t$, we get that
\begin{eqnarray}
P(0,t) &=& \E \left[ \e^{- \int_0^t r_u du} \right] \\
&=& \e^{- \int_0^t \alpha_u du} \E \left[ \e^{- \int_0^t z_u du} \right],
\end{eqnarray}
where
\begin{eqnarray}
\E \left[ \e^{I_t} \right] &=& \exp \left( \E \left[ I_t \right] + \frac{1}{2} \Var \left[ I_t \right] \right) \\
&=& \exp \left(\frac{1}{2} \Var \left[ I_t \right] \right),
\end{eqnarray}
which is consistent with Eq.~(5.28) in \cite{Pelsser2000}.

For a given scenario, i.e., $r_t(\omega) = z_t (\omega) + \alpha_t$, we get the 
time $t$ stochastic discount factor as
\begin{equation}
\e^{- \int_0^t r_u (\omega) du} = \frac{P(0,t)}{\exp \left(\frac{1}{2} \Var \left[ I_t \right] \right)} \e^{- \int_0^t z_u (\omega) du}.
\end{equation}
